---
title: "Lab 04 - Predicting MHV Change"
author: "Rachael Nicely"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

# Part 1 - Data

The following data steps are now completed in the source_file_rn.R:

    - Similar to your previous lab, create a dataset that includes 2000 and 2010 census variables drop all rural census tracts.
    - Create a variable that measures the growth of median home value from 2000 to 2010.
    - Omit cases that have a median home value less than $10,000 in 2000.
    - Omit cases with growth rates above 200%.

## Load necessary packages:

```{r, message = F, warning = F}

library( dplyr )
library( here )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( plm )
library( evaluate )
```
    

## Load wrangled datasets and prepare variables for analysis:

```{r}

# load necessary functions and objects ----
# note: all of these are R objects that will be used throughout this .rmd file
import::here("S_TYPE",
             "panel.cor",
             "panel.smooth",
             "jplot",
             "d",
             "df",
             "cbsa_stats_df",
             "mhv.00",
             # notice the use of here::here() that points to the .R file
             # where all these R objects are created
             .from = here::here("labs/wk06/source_file_rn.R"),
             .character_only = TRUE)
```

```{r}
df %>% head()
```

## Descriptives

```{r, results = 'asis'}
stargazer( df, 
           type = S_TYPE, 
           digits = 0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

## Metro Level Statistics

```{r}
# view results
cbsa_stats_df %>% head()
```


# Part 2 - Predict MHV Change

## Select at least three census variables that you feel will be good predictors of change in MHV between 2000 and 2010.

    - Percent Professionals (p.prof.00)
    - Poverty Rate (pov.rate.00)
    - Percent Unemployment (p.unemp.00)

```{r, warning = F}
# create subset to visualize in correlation matrix 
d4 <- select( d, mhv.growth, p.prof.00, pov.rate.00,  p.unemp.00 )

par( mfrow=c(3,2) )

# Unemployment
hist( d4$p.prof.00, breaks=50, col="gray20", border="white",
      yaxt="n", xlab="", ylab="", main="Percent Professionals")
hist( log(d4$p.prof.00+1), breaks=50, col="darkgreen", border="white",
      yaxt="n", xlab="", ylab="", main="Percent Professionals (logged)")
# Vacant
hist( d4$pov.rate.00, breaks=50, col="gray20", border="white",
      yaxt="n", xlab="", ylab="", main="Poverty Rate")
hist( log(d4$pov.rate.00+1), breaks=50, col="darkgreen", border="white",
      yaxt="n", xlab="", ylab="", main="Poverty Rate (logged)")
# White
hist( d4$p.unemp.00, breaks=50, col="gray20", border="white",
      yaxt="n", xlab="", ylab="", main="Percent Unemployment")
hist( log(d4$p.unemp.00+1), breaks=50, col="darkgreen", border="white",
      yaxt="n", xlab="", ylab="", main="Percent Unemployment (logged)")

```


**The distributions above in green were selected by checking for variable skew. Log transformations were used to suppress outliers for all three variables.**

## Apply log transformations:

```{r}
# reduce data density for visualization
set.seed( 1234 )

# recode some vars to remove skew
d4$p.prof.00 <- log10( d4$p.prof.00 + 1 )
d4$pov.rate.00 <- log10( d4$pov.rate.00 + 1 )
d4$p.unemp.00 <- log10( d4$p.unemp.00 + 1 )

d4.sample <- sample_n( d4, 5000 ) %>% na.omit()

# correlation plot with transformed variables
pairs( d4.sample, upper.panel=panel.cor, lower.panel=panel.smooth )
```
<br>

Check for multicollinearity:

```{r, results = 'asis'}
d.reg <- d
d.reg$mhv.growth[ d.reg$mhv.growth > 200 ] <- NA
d.reg$p.prof.00 <- log10( d.reg$p.prof.00 + 1 )
d.reg$pov.rate.00 <- log10( d.reg$pov.rate.00 + 1 )
d.reg$p.unemp.00 <- log10( d.reg$p.unemp.00 + 1 )


m1 <- lm( mhv.growth ~  pov.rate.00, data=d.reg ) # suspicious variable
m2 <- lm( mhv.growth ~  p.unemp.00, data=d.reg ) # suspicious variable
m3 <- lm( mhv.growth ~  pov.rate.00 + p.unemp.00, data=d.reg ) # test for multicollinearity

stargazer( m1, m2, m3,
           type = S_TYPE, digits = 2,
           omit.stat = c("rsq","f") )

```

<br>

**Coefficient sizes are smaller, standard errors increased, and the R-square stayed the same. These indicates that multicollinearity between the two suspicious variables, Poverty Rate and Percent Unemployment, may be a concern. This is expected considering that the variables chosen are different attributes of the same construct - neighborhood home value. **


## Run the model while including metro-level fixed effects (cbsa name or FIPS).


```{r, results = 'asis'}
model.ols <- lm( mhv.growth ~ p.prof.00 + pov.rate.00 + p.unemp.00, 
                 data = d.reg)

model.fe <- plm( mhv.growth ~ p.prof.00 + pov.rate.00 + p.unemp.00,
               index=c("cbsaname"), 
               data = d.reg)

stargazer( model.ols, model.fe, 
           type = S_TYPE, 
           dep.var.labels = ("Percent MHV Change"),
           column.labels = c("OLS", "Fixed effects"), 
           intercept.bottom = FALSE,
           omit.stat = "all", 
           digits = 2 )
```

```{r}
# intercept for each cbsa name, only show first 20 
fixef(model.fe)[1:20]

pFtest(model.fe, model.ols) 
```


<br>
<br>

## What are the results? Which factor was most important? Did it meet your expectations? Were there any variables that were not significant that you expected to be?

**According to the fixed effects results model, Percent Professional and Poverty Rate coefficients were statistically significant. Poverty Rate had the largest effect on MHV Growth, which matches our expectation since high poverty rates would indicate individuals living in poverty may not be able to afford higher valued homes and therefore affecting the the growth of home values. Percent Unemployment was not significant though it was expected to be.**   

## Explain findings to a general audience:

**In the MHV Growth fixed effect model, Percent Professional and Poverty Rate significantly affect MHV Growth. However, we have the problem of multicollinearity since the variables appear to be correlated with each other. While the variables chosen each contribute to median home value growth, Poverty Rate had the largest effect on the MHV Growth.**

<br>
<br>

