---
title: "Lab 05 - Modeling the Intervention"
author: "Group 03"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 


# Library

```{r, message = F, warning = F}
library( here )
library( import )
library( tidyverse )
library( pander )
library( dplyr )
library( knitr )
library( stargazer )
library( scales )
```

# Load Functions

```{r}
import::here("S_TYPE",
             "d",
             "df",
             "d3",
             "PLOTS",
             "%>%",
             .from = here::here("analysis/source_file_rn.R"), # here::here() points to the .R file where all these R objects are created
             .character_only = TRUE)

detach("package:import", unload = TRUE)
```

# Rodeo Dataset

```{r}
# Save data set to rodeo folder
saveRDS(d, here("data/rodeo/rodeo-data-rn.rds"))

# Load data set
readRDS(file = here("data/rodeo/rodeo-data-rn.rds"))
```

# Data Manifest

```{r}
ltdb.raw.2000s <- read.csv( here("data/raw/ltdb_std_2000_sample.csv") )
ltdb.raw.2000f <- read.csv( here("data/raw/ltdb_std_2000_fullcount.csv") )
ltdb.rodeo.2000 <- readRDS( file = here("data/rodeo/ltdb-2000.rds") )

ltdb.raw.2010s <- read.csv( here("data/raw/ltdb_std_2010_sample.csv") )
ltdb.raw.2010f <- read.csv( here("data/raw/ltdb_std_2010_fullcount.csv") )
ltdb.rodeo.2010 <- readRDS( file = here("data/rodeo/ltdb-2010.rds") )

complete.rodeo <- readRDS( file = here("data/rodeo/rodeo-data-rn.rds") )
```

```{r, eval = F}
nrow( ltdb.raw.2000s ) # 72,693
nrow( ltdb.raw.2000f ) # 72,693
nrow( ltdb.rodeo.2000 ) # 72,693

nrow( ltdb.raw.2010s ) # 73,056
nrow( ltdb.raw.2010f ) # 74,002
nrow( ltdb.rodeo.2010 ) #74,022

nrow( complete.rodeo ) # 59,066
```

# Explain why the number of rows is less

**Answer: The rodeo data were trimmed to remove rural districts and tracts with home values below $10,000 in 2000.**

# Descriptive Statistics
```{r, results = 'asis'}
stargazer( df,
           type = S_TYPE,
           digits = 0,
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

# Plots
```{r, warning = F}
### PERCENT PROFESSIONAL
gridExtra::grid.arrange( PLOTS$p_prof_2000$nmtc, 
                         PLOTS$p_prof_2000$lihtc, 
                         nrow = 1 )

### POVERTY RATES
gridExtra::grid.arrange( PLOTS$pov_rate_2000$nmtc, 
                         PLOTS$pov_rate_2000$lihtc, 
                         nrow = 1 )

### PERCENT UNEMPLOYED
gridExtra::grid.arrange( PLOTS$p_unemp_2000$nmtc, 
                         PLOTS$p_unemp_2000$lihtc, 
                         nrow = 1 )

### MHV Growth Rates (DV in Model)
gridExtra::grid.arrange( PLOTS$mhv_growth$lihtc, 
                         PLOTS$mhv_growth$nmtc, 
                         nrow = 1 )
```

# Diff-In-Diff Models

```{r, warning = F, results = 'asis'}
# set randomization seed
set.seed( 1234 )

# Model with NMTC data and control variables
m.nmtc <- lm( y ~ treat.nmtc + post + p.prof + pov.rate + p.unemp + treat.nmtc*post, data = d3 )

# Model with LIHTC data and control variables
m.lihtc <- lm( y ~ treat.lihtc + post + p.prof + pov.rate + p.unemp + treat.lihtc*post, data = d3 )

# Display model results
stargazer( m.nmtc, m.lihtc,
           type = S_TYPE, 
           column.labels = c("NMTC Model"),
           dep.var.labels = c("Median House Value (log)"),
           covariate.labels = c("Treatment - NMTC",
                                "Treatment - LIHTC",
                                "Post Time Period",
                                "Percent Professionals",
                                "Poverty Rate",
                                "Percent Unemployment",
                                "Treatment x Post - NMTC",
                                "Treatment x Post - LIHTC"),
           digits = 2 )

# NMTC Coefficients
b0.nmtc <- m.nmtc$coefficients[1] 
b1.nmtc <- m.nmtc$coefficients[2]
b2.nmtc <- m.nmtc$coefficients[3]
b3.nmtc <- m.nmtc$coefficients[4]

# LIHTC Coefficients
b0.lihtc <- m.lihtc$coefficients[1] 
b1.lihtc <- m.lihtc$coefficients[2]
b2.lihtc <- m.lihtc$coefficients[3]
b3.lihtc <- m.lihtc$coefficients[4]

# calculated post-treatment home values
T2.nmtc <- exp( b0.nmtc + b1.nmtc + b2.nmtc + b3.nmtc )
T2.lihtc <- exp( b0.lihtc + b1.lihtc + b2.lihtc + b3.lihtc )

T2.nmtc # 108074.8 
T2.lihtc # 99068.66 
```


# Are the programs effective at catalyzing neighborhood improvement?

**Answer: The New Market Tax Credit (NMTC) program is effective at catalyzing neighborhood improvement, measured by median home value (MHV) growth rate. According to the models above, the NMTC program had a larger, significant growth rate than the LIHTC program . $B_3$ (the growth rate of the treatment group above the secular growth rate) is significant for both the NMTC and LIHTC programs.**

# Reflection

How can we test the parallel lines assumption in this model? We know that growth rates change significantly between periods. The market for urban homes from 1990-2000 looks very different from the market in 2000 to 2010.

**Answer: A linear diff-in-diff approach measures the slopes from a one-unit change in the independent variable to a change in the median house value. Given that growth rates (slopes) change significantly between time periods for this evaluation, a linear approach would fail the parallel lines assumption. Instead, we use a log-linear approach where one-unit change change in the independent variable is associated with the growth rate of median house values.**  

<br>
<br>

