---
title: "Lab 03 - Descriptive Analysis"
author: "Group 03"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
---

```{r, include = F}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, warning = F, message = F}
library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( here )
library( geojsonio )  # read geoJSON map files from GitHub
library( sp )         # spatial data class sp for shapefiles
library( cartogram )  # spatial maps w/ tract size bias reduction
library( tmap )       # thematic maps
library( maptools )   # spatial object manipulation 
library( sf )         # 'simple features' flavor of shapefiles
library( tidyverse )     # data wrangling
library( tidycensus )  # pull census data
library( RColorBrewer )  # color shapefile
library( zoom )          # enlarge plot


```

```{r}
s.type <- "html"  
```

```{r}
d1 <- readRDS( here::here( "data/rodeo/LTDB-1990.rds" ) )
d2 <- readRDS( here::here( "data/rodeo/LTDB-2000.rds" ) )
md <- readRDS( here::here( "data/rodeo/LTDB-META-DATA.rds" ) )

nrow( d1 ) 
```

```{r}
import::here("compare_dfs",
             "d_sub",
             "MVH.Hist",
             "jplot",
             "mvh_compare",
             "hg",
             "gent_var",
             .from = here::here("labs/wk06/source_file_rn.R"),
             .character_only = TRUE)
```

# Part 1 - Change in Home Values
\

In urban areas, home values experienced a slight increase between the 1990-2000 period and 2000-2010. The median value increased by 0.58% which equates to a median dollar increase of $592.
\

```{r}
#merge data
d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

nrow( d )
```

```{r}
#Only Urban Tracts
d <- filter( d, urban == "urban" )
```


```{r}
#Compare shared variables
vars <- compare_dfs( df1=d1, df2=d2 )
head(vars)
```

```{r}
#Copy of dataset
d.full <- d
```


```{r}
#Create subset for the analysis
d <- d_sub(d)
head(d)
```

```{r}
#1990-2000 Inflation Rate
(1.028)^10
```

```{r}
# adjust 1990 home values for inflation 
mhv.90 <- d$mhmval90 * 1.318048  
mhv.00 <- d$mhmval00

mhv.change <- mhv.00 - mhv.90

df <- data.frame( MedianHomeValue1990=mhv.90, 
                  MedianHomeValue2000=mhv.00, 
                  Change.90.to.00=mhv.change )
```


```{r}
#Histogram change in home values
MVH.Hist()
```

```{r}
#Compare 1990 and 2000 mhv distributions
mvh_compare()
```

```{r}
#percent change
mhv.90[ mhv.90 < 10000 ] <- NA
pct.change <- mhv.change / mhv.90
summary( pct.change )
```

```{r}
# how many cases had increases above 500%
sum( pct.change > 5, na.rm=T )
```
```{r}
# preview tracts with large increases in home values 
# to see if increases make sense 

d %>% 
  filter( pct.change > 5 ) %>% 
  head()
```


```{r}
#Plot the percent change variable
hg()
```

```{r}
#Group Growth Rates By Metro Area
d$mhv.change <- mhv.change 
d$pct.change <- pct.change
d$mhv.00 <- mhv.00
d$mhv.90 <- mhv.90

d %>%
  group_by( cbsaname ) %>%
  summarize( ave.change = median( mhv.change, na.rm=T ),
             ave.change.d = dollar( round(ave.change,0) ),
             growth = 100 * median( pct.change, na.rm=T ) ) %>%
  ungroup() %>%
  arrange( - growth ) %>%
  select( - ave.change ) %>% 
  head( 25 ) %>%
  pander()
```

# Part 2 - Measuring
\

Between the 1990-2000 period and 2000-2010 there were 17,670 candidates for gentrification, and of those 871 transitioned into the advanced stages of gentrification. The candidates were determined if there was a lower than average home value and higher than average diversity. These criteria were selected as gentrification historically takes place in these poorer areas inhabited largely by marginalized racial groups. Of the candidates, the number that were determined to have transitioned into the advanced stages of gentrification experienced above average rise in home values, growth rate, and a loss of diversity. This criteria was selected as in gentrified areas home values would go from the low end of the spectrum to the higher end at a rate that would not otherwise occur in that relatively short amount of time. Growth rate would follow a similar trend to home values as development is more likely to in a less impoverished area. Loss of diversity would indicate the marginalized groups that previously lived there were displaced by whites who have become the more predominant ethnic group.   
\

```{r}
d.full$mhv.90 <- mhv.90
d.full$mhv.00 <- mhv.00
d.full$mhv.change <- mhv.change
d.full$pct.change <- pct.change
```


```{r}
d3 <- gent_var()
```


```{r, results = 'asis'}
d3 <- data.frame(d3)
stargazer( d3, 
           type=s.type, 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
# home value in lower than average home in a metro in 1990
poor.1990 <- d3$metro.mhv.pct.90 < 45  

# above average diversity for metro
diverse.1990 <- d3$metro.race.rank.90 > 45 

# home values increased more than overall city gains 
# change in percentile rank within the metro
mhv.pct.increase <- d3$metro.mhv.pct.change > 0

# faster than average growth  
# 0.58% growth in value is median for the country
home.val.rise <- d3$pct.change > 0.58 

# proportion of whites increases by more than mean 
# measured by increase in white
loss.diversity <- d3$race.change > -8

d$g.can <- poor.1990 & diverse.1990
d$g.flag <- poor.1990 & diverse.1990 & mhv.pct.increase & home.val.rise & loss.diversity

num.candidates <-  sum( poor.1990 & diverse.1990, na.rm=T )
num.gentrified <- sum( d$g.flag, na.rm=T )
```

```{r}
num.gentrified
```
```{r}
num.candidates
```

# Part 3 - Spatial Patterns
/

## Data

```{r, results = 'hide'}
# Load Crosswalk data
crosswalk <- read.csv( "https://raw.githubusercontent.com/DS4PS/cpp-529-master/master/data/cbsatocountycrosswalk.csv",  stringsAsFactors=F, colClasses="character" )

# search for city names by strings, use the ^ anchor for "begins with" 
grep( "^SAN AN", crosswalk$msaname, value=TRUE )

# Download a Shapefile with Population Data
these.sant <- crosswalk$msaname == "SAN ANTONIO, TX"
these.fips <- crosswalk$fipscounty[ these.sant ]
these.fips <- na.omit( these.fips )

state.fips <- substr( these.fips, 1, 2 )
county.fips <- substr( these.fips, 3, 5 )

# Select population data for Texas (state code 48), geometry = TRUE pulls shapefile for TX
sant.pop <- get_acs( geography = "tract", variables = "B01003_001",
           state = "48", county = county.fips[state.fips=="48"], geometry = TRUE ) %>% 
           select( GEOID, estimate ) %>%
           rename( POP = estimate )

# Load rodeo data set
census.dat <- d #census data has already been imported from source file

# Extract state codes
census.dat$GEOID <- substr( census.dat$tractid, 6, 18 )  # extract codes
census.dat$GEOID <- gsub( "-", "", census.dat$GEOID )    # remove hyphens
class( census.dat$GEOID )

# can merge an sf object and data.frame
sant <- merge( sant.pop, census.dat, by.x="GEOID", by.y="GEOID" )

# make sure there are no empty polygons
sant <- sant[ ! st_is_empty( sant ) , ]

```

## Choropleth Map Creation

```{r, warning = F, message = F, results = 'hide'}
# convert shapefile map object to a spatial version
sant.sp <- as_Spatial( sant )

# Check that it is a SpatialPolygonsDataFrame
class( sant.sp )

# project map and remove empty tracts
sant.sp <- spTransform( sant.sp, CRS("+init=epsg:3395"))
sant.sp <- sant.sp[ sant.sp$POP != 0 & (! is.na( sant.sp$POP )) , ]

# convert census tract polygons to dorling cartogram
# no idea why k=0.03 works, but it does - default is k=5
sant.sp$pop.w <- sant.sp$POP / 9000 
sant_dorling <- cartogram_dorling( x=sant.sp, weight="pop.w", k=0.05 )

# user-defined bounding box to move slocer to subjects 
bb <- st_bbox( c( xmin =  -10998490, xmax = -10890275, 
                  ymax = 3356190, ymin = 3478288 ), 
               crs = st_crs("+init=epsg:3395"))

tm_shape( sant_dorling, bbox=bb ) + 
  tm_polygons( col="mhmval00", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )
```

## Question 1: Describe the distribution of home values in 1990 - where are high and low-value tracts located in the city/cities?
  
**Answer: The highest values homes are located in the northern part of the San Antonio MSA.**
    
## Question 2: Compare values in 2000 to changes in values from 1990-2000. Do the largest gains occur in tracts with above or below-average home prices in 2000?

**Answer: In 2000 the high-value tracts expanded going further into the metro area center. The largest gains were in these high-value adjacent tracts, although there were some tracts towards the city center that saw similar gains.**  

```{r, warning = F, message = F}
#Home Value Change
tm_shape( sant_dorling, bbox=bb ) + 
  tm_polygons( col="mhv.change", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Home Value Change", title.position=c("right","top") )
```


## Question 3: Create a map that highlights tracts that are candidates for gentrification in 2000 and tracts that gentrify between 1990 and 2000. Do you find any meaningful patterns in where gentrification occurs?

**Answer: Where gentrification occurs appears to relate to a tract's proximity to tracts with higher home values. The tracts where gentrification occur tend to be on the outskirts of the city and either adjacent to or surrounded by tracts with a higher initial home value.**

```{r}
#Gentrification Candidates
tm_shape( sant_dorling, bbox=bb ) + 
  tm_polygons( col="g.can", n=10, style="cat", palette="Spectral" ) +
  tm_layout( "Gentrification Candidates", title.position=c("right","top") )
```

```{r}
#Gentrified Areas
tm_shape( sant_dorling, bbox=bb ) + 
  tm_polygons( col="g.flag", n=10, style="cat", palette="Spectral" ) +
  tm_layout( "Gentrified Areas", title.position=c("right","top") )
```
