---
title: "Lab 03 - Descriptive Analysis"
author: "Group 03"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
---

```{r, include = F}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, warning = F, message = F}
library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( here )
# set stargazer type to text for 
# previewing in RMD docs but
# convert to type HTML when knitting
# (next code chunk)

```

```{r}
s.type <- "html"  
```

```{r}
d1 <- readRDS( here::here( "data/rodeo/LTDB-1990.rds" ) )
d2 <- readRDS( here::here( "data/rodeo/LTDB-2000.rds" ) )
md <- readRDS( here::here( "data/rodeo/LTDB-META-DATA.rds" ) )

nrow( d1 ) 
```

```{r}
import::here("compare_dfs",
             "d_sub",
             "MVH.Hist",
             "jplot",
             "mvh_compare",
             "hg",
             "gent_var",
             .from = here::here("labs/wk03/utilities.R"),
             .character_only = TRUE)
```

# Part 1
\

In urban areas, home values experienced a slight increase between the 1990-2000 period and 2000-2010. The median value increased by 0.58% which equates to a median dollar increase of $592.
\

```{r}
#merge data
d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

nrow( d )
```

```{r}
#Only Urban Tracts
d <- filter( d, urban == "urban" )
```


```{r}
#Compare shared variables
vars <- compare_dfs( df1=d1, df2=d2 )
head(vars)
```

```{r}
#Copy of dataset
d.full <- d
```


```{r}
#Create subset for the analysis
d <- d_sub(d)
head(d)
```

```{r}
#1990-2000 Inflation Rate
(1.028)^10
```

```{r}
# adjust 1990 home values for inflation 
mhv.90 <- d$mhmval90 * 1.318048  
mhv.00 <- d$mhmval00

mhv.change <- mhv.00 - mhv.90

df <- data.frame( MedianHomeValue1990=mhv.90, 
                  MedianHomeValue2000=mhv.00, 
                  Change.90.to.00=mhv.change )
```


```{r}
#Histogram change in home values
MVH.Hist()
```

```{r}
#Compare 1990 and 2000 mhv distributions
mvh_compare()
```

```{r}
#percent change
mhv.90[ mhv.90 < 10000 ] <- NA
pct.change <- mhv.change / mhv.90
summary( pct.change )
```

```{r}
# how many cases had increases above 500%
sum( pct.change > 5, na.rm=T )
```
```{r}
# preview tracts with large increases in home values 
# to see if increases make sense 

d %>% 
  filter( pct.change > 5 ) %>% 
  head()
```


```{r}
#Plot the percent change variable
hg()
```

```{r}
#Group Growth Rates By Metro Area
d$mhv.change <- mhv.change 
d$pct.change <- pct.change
d$mhv.00 <- mhv.00
d$mhv.90 <- mhv.90

d %>%
  group_by( cbsaname ) %>%
  summarize( ave.change = median( mhv.change, na.rm=T ),
             ave.change.d = dollar( round(ave.change,0) ),
             growth = 100 * median( pct.change, na.rm=T ) ) %>%
  ungroup() %>%
  arrange( - growth ) %>%
  select( - ave.change ) %>% 
  head( 25 ) %>%
  pander()
```

# Part 2
\

Between the 1990-2000 period and 2000-2010 there were 17,670 candidates for gentrification, and of those 871 transitioned into the advanced stages of gentrification. The candidates were determined if there was a lower than average home value and higher than average diversity. These criteria were selected as gentrification historically takes place in these poorer areas inhabited largely by marginalized racial groups. Of the candidates, the number that were determined to have transitioned into the advanced stages of gentrification experienced above average rise in home values, growth rate, and a loss of diversity. This criteria was selected as in gentrified areas home values would go from the low end of the spectrum to the higher end at a rate that would not otherwise occur in that relatively short amount of time. Growth rate would follow a similar trend to home values as development is more likely to in a less impoverished area. Loss of diversity would indicate the marginalized groups that previously lived there were displaced by whites who have become the more predominant ethnic group.   
\

```{r}
d.full$mhv.90 <- mhv.90
d.full$mhv.00 <- mhv.00
d.full$mhv.change <- mhv.change
d.full$pct.change <- pct.change
```


```{r}
d3 <- gent_var()
```


```{r, results = 'asis'}
d3 <- data.frame(d3)
stargazer( d3, 
           type=s.type, 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
# home value in lower than average home in a metro in 1990
poor.1990 <- d3$metro.mhv.pct.90 < 45  

# above average diversity for metro
diverse.1990 <- d3$metro.race.rank.90 > 45 

# home values increased more than overall city gains 
# change in percentile rank within the metro
mhv.pct.increase <- d3$metro.mhv.pct.change > 0

# faster than average growth  
# 0.58% growth in value is median for the country
home.val.rise <- d3$pct.change > 0.58 

# proportion of whites increases by more than mean 
# measured by increase in white
loss.diversity <- d3$race.change > -8

d3$g.can <- poor.1990 & diverse.1990
d3$g.flag <- poor.1990 & diverse.1990 & mhv.pct.increase & home.val.rise & loss.diversity

num.candidates <-  sum( poor.1990 & diverse.1990, na.rm=T )
num.gentrified <- sum( d3$g.flag, na.rm=T )
```

```{r}
num.gentrified
```
```{r}
num.candidates
```

# Part 3
\

```{r}
library( geojsonio )  # read geoJSON map files from GitHub
library( sp )         # spatial data class sp for shapefiles
library( cartogram )  # spatial maps w/ tract size bias reduction
library( tmap )       # thematic maps
library( maptools )   # spatial object manipulation 
library( sf )         # 'simple features' flavor of shapefiles
```

```{r}
github.url <- "https://raw.githubusercontent.com/DS4PS/cpp-529-master/master/data/phx_dorling.geojson"
phx <- geojson_read( x=github.url,  what="sp" )
plot( phx )
```

```{r}
# create small dataframe for the merge
df <- data.frame(  tractid=d$tractid, 
        mhv.90,  mhv.00,  mhv.change,  pct.change, d3$g.can, d3$g.flag )

# create GEOID that matches GIS format

# create a geoID for merging by tract 
df$GEOID <- substr( df$tractid, 6, 18 )  # extract codes
df$GEOID <- gsub( "-", "", df$GEOID )    # remove hyphens
class( df$GEOID )
```
```{r}
nrow( phx )
```
```{r}
phx <- merge( phx, df, by.x="GEOID", by.y="GEOID" )

# make sure they have not changed or 
# you are missing rows in your data frame
# or merging with the wrong ID
nrow( phx )
```
## Question 1
\

In 1990 the high-value tracts are located on the outskirts of the city, nearly surrounding the low-value tracts located in the city center and smaller pockets moving outward. 
```{r}
#Homes Values 1990
phx <- spTransform( phx, CRS("+init=epsg:3395") )

bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                 ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395")) 

tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="mhv.90", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Homes Values 1990", title.position=c("right","top") )
```

```{r}
#Homes Values 2000
phx <- spTransform( phx, CRS("+init=epsg:3395") )

bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                 ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395")) 

tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="mhv.00", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Homes Values 2000", title.position=c("right","top") )
```

## Question 2
\

In 2000 the high-value tracts expanded to adjoining tracts. The largest gains were in these high-value adjacent tracts, although there were some tracts towards the city center that saw similar gains.   
```{r}
#Home Value Change
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="mhv.change", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Home Value Change", title.position=c("right","top") )
```

```{r}
#Home Value Percent Change
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="pct.change", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Home Value Percent Change", title.position=c("right","top") )
```

## Question 3
\

Where gentrification occurs appears to relate to a tract's proximity to tracts with higher home values. The tracts where gentrification occur tend to be on the outskirts of the city and either adjacent to or surrounded by tracts with a higher initial home value.
```{r}
#Gentrification Candidates
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="d3.g.can", n=10, style="cat", palette="Spectral" ) +
  tm_layout( "Gentrification Candidates", title.position=c("right","top") )
```

```{r}
#Gentrified Areas
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="d3.g.flag", n=10, style="cat", palette="Spectral" ) +
  tm_layout( "Gentrified Areas", title.position=c("right","top") )
```

